---
title: "Course data"
output:
  html_document: 
    css: custom_test.css
---


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%",
  message = FALSE,
  warning = FALSE
)

```



## Slides


<div class="container" style="position: relative; overflow: hidden; width: 100%; padding-top: 60%;">
<iframe class="responsive-iframe" style="position: absolute; top: 0; left: 0; right: 0; width: 100%; height: auto; aspect-ratio: 560 / 314" title="Motivation notes" src="slides/data_slides.pdf"></iframe>
</div>



## Exercise 1


Sign up for an API token and write code to download the longitudinal and cross-sectional Covid19 wastewater datasets using your token.

* [get your API token here](https://data.cdc.gov/profile/edit/developer_settings)
  * Save the value of your token somewhere, along with the email and password you used to sign up for the token! We will be using these later.
* [longitudinal data info page](https://data.cdc.gov/Public-Health-Surveillance/NWSS-Public-SARS-CoV-2-Concentration-in-Wastewater/g653-rqe2/about_data)
* [cross-sectional data info page](https://data.cdc.gov/Public-Health-Surveillance/NWSS-Public-SARS-CoV-2-Wastewater-Metric-Data/2ew6-ywp6/about_data)


<button type="button" class="collapsible">View solution</button>
<div class="content">

Download the longitudinal data, which we are storing as "covid".

```{r, eval = FALSE}
covid <- read.socrata(
  "https://data.cdc.gov/resource/g653-rqe2.json",
  app_token = "YOURAPPTOKENHERE",
  email     = "user@example.com",
  password  = "fakepassword"
)
```


Download the cross-sectional data, which we are storing as "counties".

```{r, eval = FALSE}
counties <- read.socrata(
  "https://data.cdc.gov/resource/2ew6-ywp6.json",
  app_token = "YOURAPPTOKENHERE",
  email     = "user@example.com",
  password  = "fakepassword"
)
```


</div>

## Exercise 2

For the cross-sectional data, use a query on the field `wwtp_jurisdiction` to only select rows labeled "Georgia". How many rows are returned? 



<button type="button" class="collapsible">View solution</button>
<div class="content">

```{r, eval = FALSE}
counties <- read.socrata(
  "https://data.cdc.gov/resource/2ew6-ywp6.json?wwtp_jurisdiction='Georgia'",
  app_token = "YOURAPPTOKENHERE",
  email     = "user@example.com",
  password  = "fakepassword"
)
```

Another equivalent way of writing this query is given below:

```{r, eval = FALSE}
counties <- read.socrata(
  "https://data.cdc.gov/resource/2ew6-ywp6.json?$where=wwtp_jurisdiction in('Georgia')",
  app_token = "YOURAPPTOKENHERE",
  email     = "user@example.com",
  password  = "fakepassword"
)
```


</div>







## Additional readings


* [Using SoQL queries in data pulls](https://dev.socrata.com/docs/queries/)
* [OPEN Government data act](https://data.gov/open-gov/#:~:text=The%20OPEN%20Government%20Data%20Act%20makes%20Data.gov%20a%20requirement,in%20the%20Data.gov%20catalog.)
* [Data.gov open source data](https://data.gov/)

<script language="JavaScript" src="custom_js.js"></script>

