---
title: "Git and GitHub"
output:
  html_document: 
    toc: true
    toc_float: true
    css: custom_test.css
---   

```{r, echo = FALSE, message = FALSE}
library(ggplot2)


knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  message = FALSE,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```




## Slide Deck

<div class="container" style="position: relative; overflow: hidden; width: 100%; padding-top: 100%;">
<iframe class="responsive-iframe" style="position: absolute; top: 0; left: 0; right: 0; width: 100%; height: auto; aspect-ratio: 560 / 314" title="Git notes" src="https://www.dropbox.com/scl/fi/ogqcsc7qr131vj85zutln/git.pdf?rlkey=vkdsp89q96cpui0zf5zs8ckzo&raw=1"></iframe>
</div>


<br>

## Exercises

### Initialize a repository

In this exercise, we will:

* use `git init` to initalize a git repo in your project's root directory
* view contents of the project folder before and after git initialization

Open a terminal emulator and use `cd` to navigate to the root directory of the project that you would like to turn into a git repository.

You can use the chunk of code below, where `<path/to/your/project>` should be replaced with an appropriate absolute or relative file path to the root directory of your project.

```{bash, eval = FALSE}
cd <path/to/your/project>
```

Once you have navigated to the correct directory, use the `ls` command to list files in the directory. Take note of what files are included. Then use the `ls -a` command to list __all files__ in the directory.

Now initialize a git repository using the following command.

```{bash, eval = FALSE}
git init
```

Again, use the `ls` and `ls -a` commands to list files in the directory. What has changed?

<button type="button" class="collapsible">View solution</button>
<div class="content">

The output of `ls` should be the same as before. However, when you use `ls -a` you should now see a new directory called `.git`.

The `.git` directory is where all relevant information for version control is stored. There are several files and subdirectories included in this directory. In general, you should not need to worry about what is happening in this directory. However, you should note that:

* the only thing that makes something a git repository is the presence of this `.git` directory
* deleting this directory will remove the project from version control
* if your project is also stored in a Sharepoint/Dropbox folder, you need to make sure that the contents of this directory stays synchronized across computers that work with. Strange behavior can result from incompletely synchronized `.git` folders.

If you would like to read more about what is inclued in the `.git` folder, you can check out the resources below:

* [What is in that .git directory?](https://blog.meain.io/2023/what-is-in-dot-git/)
* [The definitive deep dive into the .git folder](https://youtu.be/ADvD-DfSTSU)

</div>

<br>

### First commit

In this exercise, we will:

* use the `git add` to stage files for commit
* use `git commit` to make a commit

Before you start, make sure that your terminal emulator is still in the correct directory by using `pwd`. If you are not in the correct folder, use `cd` to navigate to the root directory of your folder.

Once you are in the correct directory, execute the following command:

```{bash, eval = FALSE}
git status
```

The output will look something like this:

````
[TODO]
````

Note that git is telling you that there are files in this directory that are not being tracked.

To make a new commit, we generally use the command

```{bash, eval=FALSE}
git add <file_name1> <file_name2> <...>
```

Use this command to stage all of the R scripts in the project for commit.

* [TODO]
* [TODO]

__DO NOT__ stage any of the following files (yet).

* [TODO]
* [TODO]

__Exercise__: What is an efficient way to use pattern matching to add all R scripts with minimal effort?

<button type="button" class="collapsible">View solution</button>
<div class="content">

We could use the following command:

```{bash, eval = FALSE}
git add *.R
```

If this command is executed from the projects root directory, then it will add all files with the `.R` extension included in your project.

Or if all your R scripts are contained in a directory, e.g., named `code`, you could add the folder.

```{bash, eval = FALSE}
git add code
```

Pattern matches can be useful for adding multiple files, but keep two things in mind:

First, you may consider avoiding commits with multiple files implementing multiple things at the same time. It may be better to split up the files across different commits.

Second, the behavior of `*` is somewhat unexpected in this circumstance:

```{bash, eval = FALSE}
git add *
```

This will add all files in your project *except* any dot-files (i.e., files whose name begins with `.`). This may miss some important files.

To add all files correctly use:

```{bash, eval = FALSE}
git add --all
```

</div>

Once you have confirmed that you have staged the appropriate files for commit, make the commit using the following command:

```{bash, eval = FALSE}
git commit -m "<add-your-message-here>"
```

Add an informative message!

Take a final look at `git status`. What has changed?

Run the following command to view the project history:

```{bash, eval = FALSE}
git log --oneline
```

<br>

### Second commit

In this exercise, we will:

* make a modification to a file
* use `git diff` to see what modifications have been made to a file
* commit a new version of the file
* use `git log` to look at the history of a project

Open [TODO] script in the R studio editor. Modify line [TODO] to read as follows:

```{r, eval = FALSE}
[TODO]
```

Save the file in R Studio. Confirm that the file has saved by ensuring that the name of the file has changed from red to black. (You would be surprised how much confusion can be caused by forgetting to save files)

Take a look at `git status`. What does it have to say about the modified file?

Use the following command to view the difference between the new version of the file and the version on the current commit.

```{bash, eval = FALSE}
git diff [TODO]
```

If all of the changes fit on one screen, the differences will simply print to the console.

If there are A LOT of changes, then you can scroll through the differences either using: enter, page up/down, or arrow keys. When you are done browsing the changes, press `q` to return to the command line.

Next, stage the file for commit:

```{bash, eval = FALSE}
git add [TODO]
```

__Exercise__: We are now going to do something a little bit scary. We are going to run the `git commit` command but forget to include a message 😱

```{bash, eval = FALSE}
git commit
```

*More than likely* you will see the following on the screen:

````
# Please enter the commit message for your changes. Lines starting
# with '#' will be ignored, and an empty message aborts the commit.
````

But if you try to type nothing happens. What the...

The exercise is simple: escape!

<button type="button" class="collapsible">View solution</button>
<div class="content">

What happened here? When you forget to include a commit message, git launches a text editor program on your computer. The default text editor that git uses is [`vim`](https://vim.org), a powerful, but painfully minimal and more painfully complex editor.

The key part of this solution is: DON'T PANIC. We will get through this together.

Follow these precise steps:

* type `i`
  * you should see `-- INSERT --` at the bottom of the screen
  * now when you type you will actually see letters appear!
* use the keyboard as normal to type your commit message
  * notice that you cannot use the mouse to select the cursors position, you have to use the arrow keys
* when you have finished your message press `esc`
  * you should no longer see `-- INSERT --` at the bottom
* type `:x` (colon and the x key)
  * you will see `:x` appear at the bottom of the screen
* press enter to save the commit message and quit `vim`

This will complete the commit. But what the heck just happened?!

In vim, we use `i` to enter `INSERT` mode, which is where we actually get to type things. 

Pressing `esc` exited from `INSERT` mode (back to what is called `NORMAL` mode). 

The colon key takes us into command-line mode. This is where we can execute a host of commands (e.g., find and replace text, save a document, etc...). Think of this like a text-based tool bar for `vim`.

In command-line mode, we execute the command `x` which is to save and close the document.

This is exactly what git was asking us to do to complete the commit -- save a message.

If all this was too scary for you, you have two options:

Option 1: abandon the commit! Before you do anything else in `vim` type `:q!`, which will quit vim without saving. This abandons the commit. Then you can use the one-liner `git commit -m "<with-a-message-this-time-dummy>"` to complete the commit.

Option 2: configure a different text editor for git! Read more [here](https://docs.github.com/en/get-started/getting-started-with-git/associating-text-editors-with-git).

</div>


<br>

### Third commit

In this exercise, we will:

* add a temporary file to a new commit
* demonstrate renaming files using `git mv` in a new commit
* delete a file using `git rm` in a new commit

In R Studio, open a new file, add whatever you want to the file, and save it in the [TODO] directory. Name the file `foo.R`.

Use the following commands to commit this new file:

```{bash, eval = FALSE}
git add [TODO]/foo.R
git commit -m "Add temporary file to demonstrate renaming and removing"
``` 

To rename the file use:

```{bash, eval = FALSE}
git mv [TODO]/foo.R [TODO]/bar.R
``` 

Check `git status` to see what git thinks about this change! To commit, run:

```{bash, eval = FALSE}
git commit -m "Rename temporary file"
```

To remove the file use:

```{bash, eval = FALSE}
git rm [TODO]/bar.R
```

Check `git status` to see what git thinks about this change! To commit, run:

```{bash, eval = FALSE}
git commit -m "Remove temporary file"
```

__Exercise__: Repeat the above process of creating a new file in the `[TODO]` directory named `foo.R`. Add it to a new commit using `git add` and `git commit`. Now suppose you forget to use `git mv` to rename the file and instead use plain old `mv`:

```{bash, eval = FALSE}
mv [TODO]/foo.R [TODO]/bar.R
```

Check `git status` to see what git thinks has happened. What do you need to do to complete the process of committing the renamed file? 

<button type="button" class="collapsible">View solution</button>
<div class="content">

Notice that `git status` should show `[TODO/foo.R]` as being deleted and show `[TODO/bar.R]` as an Untracked file.

To complete the commit we need to run:

```{bash, eval = FALSE}
git add [TODO/bar.R]
git rm [TODO/foo.R]
git commit -m "Rename temporary file"
```

</div>

<br>

__Exercise__: In the previous exercise, you again renamed the file `[TODO]/foo.R` to `[TODO/bar.R]`. Now create a commit that removes `[TODO/bar.R]` but __does not delete__ the file from your local directory.

<button type="button" class="collapsible">View solution</button>
<div class="content">

This can be accomplished using the following (not very intuitively named) option:

```{bash, eval = FALSE}
git remove --cached [TODO]/bar.R
git commit -m "Remove temporary file"
```

Using `ls` should confirm that `[TODO]/bar.R` has not been deleted from your directory. 

Checking `git status` should show that `[TODO]/bar.R` is now treated as any other untracked file. For example, it could be added back to a new commit.

</div>

<br>

### Ignoring files

In this exercise, we will:

* create a `.gitignore` file to ignore certain files

After completion of the previous exercise, you should have a file named `[TODO]/bar.R` that is untracked by git. If you did not complete the previous exercise or do not have such a file available, then create one by running:

```{bash, eval = FALSE}
touch [TODO]/bar.R
```

Now we are ready to create our `.gitignore` file. In R Studio, select File: New File: Text file.

Save the file as `.gitignore` by clicking File: Save as: and type `.gitignore`.

Now add the following text to the `.gitignore` file

````
# don't want credentials in git!
.Renviron
# ignoring this file to see behavior
[TODO]/bar.R
````

Save the file when you are done adding the text.

Now check `git status` again. You should no longer see [TODO]/bar.R appear as an untracked file.

__Exercise__: You may have also noticed that git is now treating the `.gitignore` file itself as an untracked file. Here's what should be an easy exercise by now (hopefully!) -- make a new commit that includes the `.gitignore` file

<button type="button" class="collapsible">View solution</button>
<div class="content">

We can execute the following commands:

```{bash, eval = FALSE}
git add .gitignore
git commit -m "Add .gitignore to repo"
```

</div>

<br>

### Pushing files to GitHub

In this exercise, we will:

* create an empty GitHub repository
* use `git remote add` to create a link between our local repository and the GitHub repository
* create ssh credentials to allow us to push to GitHub
* use `git push` to push files to GitHub

Follow these instructions to create an __empty__ GitHub repository:

* From GitHub dashboard, click on the `+` symbol in the upper right-hand corner and select New repository
* Give the repository a name
* __DO NOT__ check the `Add README file` box
* __DO NOT__ add a `.gitignore`
* __DO NOT__ choose a license
* Click Create repository

Now go back to your R Studio terminal emulator and run the following command to add the GitHub repository as a remote named `origin` to your local repository:

```{bash, eval = FALSE}
git remote add origin git@github.com:<github_user>/<repo_name>
```

* `<github_user>` = your GitHub username
* `<repo_name>` = the name of your GitHub repository

Confirm that the remote was added as expected by listing the remotes for your repository:

```{bash, eval = FALSE}
git remote –v
```

Now we are ready to push, except that we need to prove to GitHub that we have permission to push to a repository that we own. To do this we will use [private/public keys](https://crypto.stackexchange.com/questions/30462/private-and-public-keys).

First, check if any keys exist already:

```{bash, eval = FALSE}
ls -al ~/.ssh
```

Look for files named

* `id_rsa.pub`
* `id_ecdsa.pub`
* `id_ed25519.pub`

If you do not have these files or if the `.ssh` directory does not exist, then you need to create a key pair.

To create a key pair, run the following command:

```{bash, eval = FALSE}
ssh-keygen -t ed25519 -C "<github_account@email.com>"
```

Be sure to replace `<github_account@email.com>` with the email address you used to create your GitHub account.

When prompted to `"Enter a file in which to save the key"`, press `Enter` to accept the default file location.

For simplicity, when prompted to enter a passphrase, just press `Enter` twice to use no passphrase. If you would like the additional security of a password, read more about [configuring authentication agents](https://docs.github.com/en/authentication/connecting-to-github-with-ssh/working-with-ssh-key-passphrases).

Re-run the `ls --al ~/.ssh` command to confirm the key was created successfully. You should see a file with `.pub` ending displayed.

Now head back to your web browser and GitHub. Navigate to the GitHub dashboard (i.e., your landing page once you sign in). 

* Click on the icon in the top right corner of the page (this will be your picture, if you've uploaded one to your profile, otherwise it will be a pixelated image). 

* Select Settings from the dropdown menu.

* When the settings menu opens, on the left side of the screen, click "SSH and GPG keys".

* Click the green button "New SSH Key"

* Give the key a title (can be arbitrary)

* Open the file `~/.ssh/id_<something>.pub` in R Studio.
  * Recall that `~` means your `HOME` directory. To remind yourself of what this directory is you can run `echo $HOME` in the terminal.
  * You may need to ensure that hidden files are shown in your file finder to see this file in order to open in in R Studio.

* Copy the contents of the file to the Key box on GitHub

* Click Add SSH key

* Test your connection by executing ssh -T git@github.com

You should see a message that reads:

````
Hi <user>! You've successfully authenticated, but GitHub does not provide shell access.
````

Now we are finally ready to push to GitHub. Use the command: 

```{bash, eval = FALSE}
git push origin main
```

If you get an error about `no branch named main`, then try:

```{bash, eval = FALSE}
git push origin master
```

Once you have pushed, refresh your browser on your GitHub repository. You should be able to see your code!

<br>

### Practicing add/commit/push

In this exercise, we will:

* write a `README.md` file for your GitHub repository
* add, commit, and push `README.md` to your GitHub repository

__Exercise__: Create a file called `README.md` using R Studio. Put the following contents into the file:

````
# COVID-19 Wastewater Analysis

This repo houses test code from the SISMID 2024 course Hitchhikers Guide to reproducibility.
````

Make a new commit that includes the `README.md` and then push the commit to GitHub.

<button type="button" class="collapsible">View solution</button>
<div class="content">

To add and commit the file:

```{bash, eval = FALSE}
git add README.md
git commit -m "Add README to repo"
```

To push to GitHub:

```{bash, eval = FALSE}
# replace main with master as appropriate
git push origin main
```

</div>

<br> 

__Exercise__: Instead of creating a markdown file directly, we will instead use R Markdown to create the file. 

Create a new file called `README.Rmd`. In the `yaml` header, include `output_type: github_document`. Write a new `README` file using R Markdown syntax. Feel free to include code chunks or whatever. This is your opportunity to tell people how your code works!

Knit the document, outputting a modified version of `README.md`. Create a commit that includes both the source code `README.Rmd` and the updated `README.md` file.

<button type="button" class="collapsible">View solution</button>
<div class="content">

In the `README.Rmd` file, we might include the following contents.

````
---
title: COVID-19 Wastewater analysis
output_type: github_document
---

This repo houses test code from the SISMID 2024 course Hitchhikers Guide to reproducibility.

I am including some R code just to demonstrate that R Markdown helps with nice README formatting.

`r ''````{r}
a <- 10
b <- 10
a + b
```
````

Please be sure to include the line `This repo houses test code from the SISMID 2024 course Hitchhikers Guide to reproducibility.` It is referenced in another exercise below.
`
Knit the document to create README.md. Then create a commit and push it using:

```{bash, eval = FALSE}
git add README.Rmd README.md
git commit -m "Update README using R Markdown"
git push origin main
```

Refresh your GitHub repository to see the changes to your `README`.

</div>

<br>

### Creating and merging branches

In this series of exercises, we will:

* create a new branch
* make a commit on the new branch
* merge the new branch with `main`
* delete the new branch

__Exercise__: Create a branch named `plot`. Checkout this branch and begin to make minor modifications to the figure generated by `[TODO]/[TODO].R`. Create a commit with the new changes.

<button type="button" class="collapsible">View solution</button>
<div class="content">

To create a new branch we run:

```{bash, eval = FALSE}
git branch plot
```

To move the `HEAD` pointer to the new branch we run:

```{bash, eval = FALSE}
git checkout plot
```

Note that there is an option that can be used as a shortcut to do both of these steps at the same time! 

```{bash, eval = FALSE}
git checkout -b plot
```

Now you can modify the plotting file as you please. E.g., you might change the axis labels. Save the file when you are done.

Make a commit on the `plot` branch:

```{bash, eval = FALSE}
git add [TODO]/[TODO.R]
git commit -m "Update plotting code"
```

Use `git log --oneline` to confirm that `plot` branch is now 1 commit ahead of `main`.

</div>

Now that we have made a commit on the `plot` branch, let's pause to see how we would view the older version of the code that is still on the `main` branch.

To do this, we simply move the `HEAD` pointer by checking out `main`.

```{bash, eval = FALSE}
git checkout main
```

View the contents of `[TODO]/[TODO.R]` in R Studio. You should see the older version of the code.


__Exercise__: Merge `plot` into `main`.

<button type="button" class="collapsible">View solution</button>
<div class="content">

Be sure that you have `main` checked out. If you are not sure, use `git branch` to see. The branch with the asterisk is your current branch. If you are not on `main`, use `git checkout main` to switch branches.

Once you are satisfied that you are on the `main` branch run:

```{bash, eval = FALSE}
git merge plot
```

Check `git log --oneline` to confirm that that `main` pointer has appropriately moved. 

</div>

<br>

### Resolving merge conflicts

In this exercise, we will:

* create a new branch
* make a commit on the new branch
* checkout the `main` branch
* make a commit on the `main` branch
* attempt to merge when conflicts are present
* resolve conflicts and complete the merge

We are going to create a new branch specifically to work on our `README` again. Let's call the branch `readme`:

```{bash, eval = FALSE}
git branch readme
git checkout readme
```

Open `README.Rmd` in R Studio. Modify the line:

````
This repo houses test code from the SISMID 2024 course Hitchhikers Guide to reproducibility.
````

to read

````
This repo houses test code from the *FANTASTIC* SISMID 2024 course Hitchhikers Guide to reproducibility.
````

Save the file, knit it, and commit the changes:

```{bash, eval = FALSE}
# use pattern matching to add both files
git add README*
git commit -m "Update READMEs with thoughts on course."
```

Now checkout the `main` branch. When you open `README.Rmd` you should see that the text has reverted to

````
This repo houses test code from the SISMID 2024 course Hitchhikers Guide to reproducibility.
````

Change it to

````
This repo houses test code from the *absolutely miserable* SISMID 2024 course Hitchhikers Guide to reproducibility.
````

Save the file, knit it, and commit the changes:

```{bash, eval = FALSE}
# use pattern matching to add both files
git add README*
git commit -m "Update READMEs with new thoughts on course."
```

__Exercise__: Attempt to merge `readme` into `main`. Identify the merge conflicts by selecting your true feelings for this course. Commit the result.

<button type="button" class="collapsible">View solution</button>
<div class="content">

There will be conflicts in `README.Rmd` and `README.md`. In both files you should see something like this:

````
<<<<<<< HEAD
This repo houses test code from the *absolutely miserable* SISMID 2024 course Hitchhikers Guide to reproducibility.
=======
This repo houses test code from the *FANTASTIC* SISMID 2024 course Hitchhikers Guide to reproducibility.
>>>>>>> readme
````

Remember that we are on the `main` branch so `HEAD` is currently point to `main`. Thus, the text that appears first is that from the `main` branch; the text on the bottom is from `readme`.

Express your true feelings by removing the text that you do not want to keep (and the extra symbols). Save the file and commit the result:

```{bash, eval = FALSE}
git add README*
git commit -m "Resolve my conflict(ed feelings)"
```

</div>

<br>

### Pull requests and upstream tracking

This is a longer exercising involving a partner. Designate one partner as User A and the other as User B.

First, User B will:

* fork User A's existing repository
* `clone` the repository to create a local repository
* make changes to the local repository
* `push` changes to GitHub
* submit a pull request

Next, User A will:

* add User B's repository as a remote
* `fetch` User B's repository
* merge User B's changes to their local repository
* push their changes to GitHub

Then, we will use `upstream` tracking by User B of User A's repository. User A will:

* make changes to their local repository
* `push` those changes to GitHub

Finally, User B will:

* add User A's repository as an `upstream` remote
* `fetch` User A's repository
* merge User A's changes into their local repository
* push their changes to GitHub

Specific instructions are given in the subheadings below.

#### User A: Confirm GitHub repository is accessible

*User A* should have a GitHub repository created associated with the exercises above. Share your GitHub user name and your GitHub repository name with *User A*.

#### User B: Fork and clone the repository

*User B* should now fork and clone *User A*'s repository on GitHub using the following steps.

To do this, *User B* should navigate to `https://github.com/<user_a_name>/<user_a_repo>` and click "Fork" to create a fork of *User A*'s repository.

* replace `<user_a_name>` with *User A*'s GitHub user name
* replace `<user_a_repo>` with *User A*'s GitHub repository name

Recall that this creates a copy of *User A*'s GitHub repository on *User B*'s GitHub. This repository is now viewable at `https://github.com/<user_b_name>/<user_b_repo>`.

*User B* is now going to clone __their copy__ (not User A's copy!) of the GitHub repository.

To do this, *User B* should use `cd` in their terminal to navigate to a directory where they wish to download *User A*'s repository.

*User B* should execute 

```{bash, eval = FALSE}
# make sure this is User B's repo
# and NOT User A's!!!
git clone git@github.com:<user_b_name>/<user_b_repo>` 
```

* this will clone *User B*'s fork of the repository
* be sure to use the `git@github.com:<user_b_name>/<user_b_repo>` syntax and __not__ `https://github.com/<user_b_name>/<user_b_repo>` syntax.
  * You can confirm what web address was used to add the remote by executing `git remote -v`.

If the output of `git remote -v` shows that you accidentally used `https://` syntax in your `git clone` command then *User B* should remove the origin `remote` using `git remote remove origin` and then re-add the remote using "ssh"-style syntax: `git remote add origin git@github.com:<user_b_name>/<user_b_repo>`.

*User B* should confirm that a folder called `<user_b_repo>` was added to the current working directory of their terminal.

#### User B: Update the repository and submit a pull request

*User B* will now make a new branch and make updates to *User A*'s repo on that branch. *User B* should complete the following steps:

* Create and checkout a new branch called `feature` by executing `git checkout -b feature`
  * recall that this simultaneously creates and checks out a new branch called `feature`
  * confirm that *User B* has switched to the new branch by executing `git branch`. You should see a star next to `feature`

* Modify something about the `[TODO]/[TODO.R]` script. E.g., [TODO]

* After completing changes, appropriately use `git add` and `git commit` to make a new commit along the `feature` branch

* Push the `feature` branch to GitHub.

```{bash, eval = FALSE}
git push origin feature
```

* Submit a pull request to *User A*'s repository.
  * the pull request should request that *User B*'s `feature` branch be merged into *User A*'s `main` branch.

#### User A: test out pull request code

*User A* will now `fetch` the code submitted by *User B*, test it out, and eventually merge it into their `main` branch, thereby closing the pull request. To do this *User A* should 

* Add *User B*'s repository as a remote named `userb`.
  
```{bash, eval = FALSE}
git remote add userb git@github.com:<user_b_name>/<user_b_repo>`
```
    * replace `<user_b_name>` with *User B*'s GitHub user name
    * replace `<user_b_repo>` with *User B*'s GitHub repository name

* `fetch` from *User B*'s repository.

```{bash, eval = FALSE}
git fetch <user_b_remote_name>
```

  * Recall that this command downloads the contents of *User B*'s repository and creates a remote tracking branch called `userb/feature`.

* *User A* should checkout the remote tracking branch to try out the new code

```{bash, eval = FALSE}
git checkout userb/feature 
``` 
  
* *User A* should test out the code on the `feature` branch.
  - E.g., confirm that the report builds properly with the updated code

* When *User A* is satisfied that the code works as expected, they should merge the `userb/feature` branch into `main`.
```{bash, eval = FALSE}
git checkout main
git merge userb/feature
```

* *User A* should push the updated `main` branch to GitHub.

```{bash, eval = FALSE}
git push origin main
```
Both users should now see *User B*'s pull request as "merged" on GitHub.

#### User A: New commits

*User A* will now create new commits on `main` that *User B* will need to track in order to keep up with the current state of the repository.

* *User A* should add some additional (trivial) code to the code base.
  * If you can't think of anything interesting to do, just add a comment to an R script.

* *User A* should appropriately `add` and `commit` this changes.

* *User A* should appropriately `push` these changes to their GitHub repository.


#### User B update local repository

Now *User B* wishes to merge the changes made by *User A* into their `main` branch, so that their repository stays up to date with *User A*'s. 

* *User B* should add a remote called `upstream` that points to *User A*'s GitHub repository.

```{bash, eval = FALSE}
git remote add upstream git@github.com <user_a_name>/<user_a_repo>`
```

User B* should now `fetch` from the `upstream` remote and merge `upstream/main` into `main`

* To download the current contents of *User A*'s repository (i.e., the `upstream` remote) to *User B*'s computer use:

```{bash, eval = FALSE}
git fetch upstream
```

* *User B* should ensure that they are on their own `main` branch by checking `git branch`. If they are not on `main`, use  `git checkout main`

* Merge *User A*'s changes into *User B*'s `main` branch.

```{bash, eval = FALSE}
git merge upstream/main
```
* Confirm that *User B*'s local files on the `main` branch now look like *User A*'s files from GitHub.

* Finally, *User B* should `push` their `main` branch to their GitHub fork to make sure the fork is up to date as well:

```{bash, eval = FALSE}
git push origin main
```

If you have time, reverse roles of User A and B and repeat the exercise. 


<script>
  var coll = document.getElementsByClassName("collapsible");
var i;

for (i = 0; i < coll.length; i++) {
  coll[i].addEventListener("click", function() {
    this.classList.toggle("active");
    var content = this.nextElementSibling;
    if (content.style.display === "block") {
      content.style.display = "none";
    } else {
      content.style.display = "block";
    }
  });
}
</script>